---
title: "Univalle - Organizacion Industrial, 2017"
author: "Trabajo 1"
date: "20 de diciembre de 2017"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

This problem set requires that you generate simulated data of a market with 100 differentiated products
with one observed characteristic, besides the price. The characteristic of product $j \in \{1,2,...,100 \}$ is $x_j=j$. You are required to generate the unobserved product attributes which are assumed to be unknown in the estimation. You'll also be able to generate as many instruments as you need. In the first part, a model with homogeneous consumers is generated and estimated. In the second part a
more complex model with heterogeneous consumers is generated and estimated. You will have to use a mathematical software such as Matlab, R, Gauss, Fortran, etc. Each student must hand a (different) typed homework. The homework should be written in the style of a paper, explaining what is done in each section, writing down the equations, etc.

1. Generate prices, market shares and unobserved characteristics of the 100 products in the market based on the following assumptions:

$$u_{ij}=\delta_j+\epsilon_{ij}$$
 
where $\delta_j = \beta x_j -\alpha p_j + \xi_j$ is the mean utility level obtained by consumer $i$ when purchasing product $j$ and $\epsilon_ij$ is a type-I extreme value disturbance.

Generate 
$$\xi_j \sim U(0,1)$$

Generate prices using the following equation

$$p_j = \eta x_j + \xi_j + \sum_{\iota=1...L} \varepsilon_{\iota,j}+\omega_j$$

where 
$$\varepsilon_{\iota=1...L} \overset{_{iid}}{\sim} N(0,1)$$
$$\omega_j\overset{_{iid}}{\sim} N(0,1)$$

```{r}
library(evd) #generates the required T1EV draws

ni <- 10000 #number of consumers
nj <- 100 #number of differentiated products
nz <- 5 #number of intruments
beta <- 1 #made up parameter on characteristics
alpha <- 1 #made up parameter in price
eta <- 1 #made up parameter on characteristic (price eq)
xj <- data.frame(c(1:nj)) #observed non-price characteristic

#UTILITY
set.seed(85719)
pigmaj <- runif(100,0,1) #unobserved characteristics of good j

epsilonij <- data.frame(matrix(nrow = ni,ncol = nj)) #empty dataframe for individual shocks
colnames(epsilonij) <- paste0('good',c(1:100))

ptm <- proc.time()
for(i in 1:ni){
  epsilonij[i,] <- rgumbel(nj,0,1) #individual utility shock T1EV

}
proc.time()-ptm


    #PRICES
    
    omegaj <- rnorm(100,0,1) #disturbance price equation of good j
    
    #instruments
    vareps <- data.frame(matrix(nrow = nj,ncol = nz))
    for (i in 1:nz){
      vareps[,i] <- rnorm(nj,0,1)
    }
    colnames(vareps) <- paste('z', 1:5)
    
    #creates prices
    priceFun <- function(xj,pigma,instr, shock) eta*xj + pigma + instr + shock
    
    varepsum <- rowSums(vareps)
    
    priceVec <- priceFun(xj,pigmaj,varepsum,omegaj)



#creates mean utilities for the nj differentiated products
deltaFun <- function(xj,a,price,pig) beta*xj - a*price + pig
meanUVec <- deltaFun(xj,alpha,priceVec,pigmaj)


# creates utilities from a vector of mean utilities (nj) j and an n-vector of unobserved random attributes (ni)
uVec <- data.frame(matrix(,nrow = ni,ncol = nj))

ptm <- proc.time()
for(i in 1:ni) {
  uVec[i,] <- epsilonij[i,] + t(meanUVec)
}

proc.time()-ptm

uOg <- 0 #utility outside good

uVec1 <- cbind(uVec,uOg) #adds column wiht utility of outside good

discr <- apply(uVec1,1,which.max) #goods that yield highest utility
discrUs <- apply(uVec1,1,max)

choicesGoods <- data.frame(names(uVec1)[discr]) #same but names

prop.table(table(choicesGoods)) #market shares




```



